---
title: "DATA 605 Week 2 Discussion"
author: "Ilya Kats"
date: "September 6, 2017"
params:
  username: "root"
  password: ""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

## Chapter D Section PDM Exercise C30

Each of the archetypes below is a system of equations with a square coefficient matrix, or is a square matrix itself. Compute the determinant of each matrix, noting how Theorem SMZD indicates when the matrix is singular or nonsingular.

Archetype A, Archetype B, Archetype F, Archetype K, Archetype L

#### Archetype A

$$
\begin{bmatrix}
  1 & -1 & 2 \\
  2 & 1 & 1 \\
  1 & 1 & 0
\end{bmatrix}
$$

Calculate determinant by expanding on the third row.

$$
\begin{split}
det\ A & = 1
\begin{bmatrix}
  -1 & 2 \\
  1 & 1 \\
\end{bmatrix}
-1
\begin{bmatrix}
  1 & 2 \\
  2 & 1 \\  
\end{bmatrix}
+0
\begin{bmatrix}
  1 & -1 \\
  2 & 1 \\
\end{bmatrix} \\
&= 1 (-1-2) - 1 (1-4) + 0 \\
&= -3 + 3 + 0 \\
&= 0
\end{split}
$$

```{r}
A <- matrix(c(1,2,1,-1,1,1,2,1,0), nrow=3)
A
det(A)
```

Since $det\ A = 0$, per Theorem SMZD $A$ is singular.

Calculate all other determinants using R.

```{r}
B <- matrix(c(-7,5,1,-6,5,0,-12,7,4), nrow=3)
F <- matrix(c(33,99,78,-9,-16,-47,-36,2,10,27,17,3,-2,-7,-6,4), nrow=4)
K <- matrix(c(10,12,-30,27,18,18,-2,-21,30,24,24,-6,-23,36,30,24,0,-30,37,30,-12,-18,39,-30,-20), nrow=5)
L <- matrix(c(-2,-6,10,-7,-4,-1,-5,7,-5,-3,-2,-4,7,-6,-4,-4,-4,10,-9,-6,4,6,-13,10,6), nrow=5)
```

#### Archetype B

```{r}
B
det(B)
```

Since $det\ B \ne 0$, per Theorem SMZD $B$ is nonsingular.

#### Archetype F

```{r}
F
det(F)
```

Since $det\ F \ne 0$, per Theorem SMZD $F$ is nonsingular.

#### Archetype K

```{r}
K
det(K)
```

Since $det\ K \ne 0$, per Theorem SMZD $K$ is nonsingular.

#### Archetype L

```{r}
L
det(L)
```

R calculates $det\ L$ as very close to zero, but not exactly zero. I have a suspicion that this is just a rounding error. Check if inverse matrix exists. 

```{r}
library(matlib)
inv(L)
```

$L$ is not inversible. Therefore, it is singular. Therefore, $det\ L = 0$.