---
title: "DATA 605 Week 1 Homework"
author: "Ilya Kats"
date: "September 2, 2017"
params:
  username: "root"
  password: ""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Set 1

My approach for problem set 1 was to do all computations manually, but then I have double-checked them with R code. *Side note:* It took me embarrassingly long time to realize that I need to make sure that the aspect ratios of x axis and y axis in the plot should match if I want to double-check the angle visually.

### Part (1)

If

$u=[0.5, 0.5]$

$v=[3, -4]$

Then dot product is

$$
\begin{split}
  u \cdot v &= [0.5, 0.5] \cdot [3, -4] \\
    &= 0.5 * 3 + 0.5 * (-4) \\
    &= 1.5 - 2 \\
    &= -0.5 
\end{split}
$$

Check with R...

```{r}
u <- c(0.5, 0.5)
v <- c(3, -4)
dot_uv <- drop(u %*% v)
dot_uv
```

### Part (2)

Length of vector $u$

$$
\begin{split}
  \left\| u \right\| &= \sqrt{0.5^2 + 0.5^2} \\
  &= \sqrt{0.25 + 0.25} \\
  &= \sqrt{0.5} \approx{0.7071}
\end{split}
$$

Length of vector $v$

$$
\begin{split}
  \left\| v \right\| &= \sqrt{3^2 + {(-4)}^2} \\
  &= \sqrt{9 + 16} \\
  &= \sqrt{25} \\
  &= 5
\end{split}
$$

Check with R...

```{r}
len_u <- sqrt(u[1]^2 + u[2]^2)
len_v <- sqrt(v[1]^2 + v[2]^2)
len_u
len_v
```

### Part (3)

$$
\begin{split}
  3u - 2v &= 3 [0.5, 0.5] - 2 [3, -4] \\
  &= [3*0.5, 3*0.5] - [2*3, 2*(-4)] \\
  &= [1.5, 1.5] - [6, -8] \\
  &= [1.5 - 6, 1.5 - (-8)] \\
  &= [-4.5, 9.5]
\end{split}
$$

Check with R...

```{r}
3*u - 2*v
```

### Part (4)

Let $\theta$ be an angle between vectors $u$ and $v$, then 

$$
\begin{split}
  \cos(\theta) &= \frac{u \cdot v}{\left\| u \right\|\left\| v \right\|}\\
  &= \frac{-0.5}{0.7071 * 5} \approx -0.1414
\end{split}
$$

```{r}
cos_theta <- dot_uv / (len_u * len_v)
cos_theta

theta <- acos(cos_theta) * 180/pi
theta
```

Based on R code above the angle is $\theta = cos^{-1}(-0.1414) \approx 98.13^{\circ}$.

Confirm results by plotting two vectors...

```{r}
plot(NA, xlim=c(0,3), ylim=c(-4,0.5), xlab="X", ylab="Y", asp=1)
arrows(0,0,0.5,0.5)
arrows(0,0,3,-4)
grid()
```

## Problem Set 2

### Generate Row-Equivalent Matrix in Echelon Form

The following code is based on the **Gauss-Jordan elimination** algorithm presented in Subsection RREF of the  *A First Course in Linear Algebra* book by Robert A. Beezer. The `echelon` function works on any size coefficient matrix.

```{r}
# Generate row-equivalent matrix in echelon form
# A is a coefficient matrix
# b is vector of constants
echelon <- function(A, b) {
  A1 <- cbind(A,matrix(b))

  m <- dim(A)[1]
  n <- dim(A)[2]

  r <- 0
  for(j in 1:n) {
    i <- r+1
    if(i <= m) {
      A1ij <- A1[i,j]
    }
    while (i <= m & A1ij==0) {
      i <- i+1
      if(i <= m) {
        A1ij <- A1[i,j]
      }
    }
    if(i < m+1) {
      r <- r+1
      
      # Swapping rows i and r (row operation 1)
      temprow <- A1[i, ]
      A1[i, ] <- A1[r, ]
      A1[r, ] <- temprow
      
      # Scale A[r,j] to 1 (row operation 2)
      multiplier <- A1[r,j]
      for(t in 1:(n+1)) {
        A1[r,t] <- A1[r,t] / multiplier 
      }
      
      # Zero out all other values in the column (row operation 3)
      for(k in 1:m) {
        if(k!=r) {
          multiplier <- A1[k,j]
          for(t in 1:(n+1)) {
            A1[k,t] <- A1[k,t] - A1[r,t] * multiplier
          }
        }
      }
    }
  }
  return(A1)
}
```

### Examples

Even though the code above works on any size system of equations, examples and code below only test systems of 3 variables and 3 constraints. The code in all examples is the same. To test other examples, only adjustments to `A` and `b` are needed.

#### EXAMPLE 1

```{r}
# Coefficient matrix
A <- matrix(c(1,2,-1,1,-1,-2,3,5,4), nrow = 3)
A

# Vector of constants
b <- c(1,2,6)
b

# Row-equivalent matrix in echelon form
Ae <- echelon(A, b)
Ae

# Find and output solution(s)
onesolution <- TRUE
for(i in 1:dim(A)[1]) {
  if (all(Ae[i, 1:dim(A)[2]]==0)) {
    if (Ae[i,dim(Ae)[2]]==0) {
      msg <- "There are infinitely many solutions."
    } else {
      msg <- "There are no solutions."
    }
    onesolution <- FALSE
  }
}
```

**SOLUTION:**

```{r}
if (onesolution) {
  msg <- "There is only one solution."
  Ae[,dim(Ae)[2]]
}

print(msg)
```

Because of rounding when converting the systen to echelon form, the solution does not match exactly the solution specified in the problem ([-1.55, -0.32, 0.95]).

#### EXAMPLE 2

```{r}
# Coefficient matrix
A <- matrix(c(3,5,2,4,-2,-2,-1,1,1), nrow = 3)
A

# Vector of constants
b <- c(8,4,1)
b

# Row-equivalent matrix in echelon form
Ae <- echelon(A, b)
Ae

# Find and output solution(s)
onesolution <- TRUE
for(i in 1:dim(A)[1]) {
  if (all(Ae[i, 1:dim(A)[2]]==0)) {
    if (Ae[i,dim(Ae)[2]]==0) {
      msg <- "There are infinitely many solutions."
    } else {
      msg <- "There are no solutions."
    }
    onesolution <- FALSE
  }
}
```

**SOLUTION:**

```{r}
if (onesolution) {
  msg <- "There is only one solution."
  Ae[,dim(Ae)[2]]
}

print(msg)
```

#### EXAMPLE 3

```{r}
# Coefficient matrix
A <- matrix(c(5,4,-1,-2,-4,1,1,-8,2), nrow = 3)
A

# Vector of constants
b <- c(3,2,-3)
b

# Row-equivalent matrix in echelon form
Ae <- echelon(A, b)
Ae

# Find and output solution(s)
onesolution <- TRUE
for(i in 1:dim(A)[1]) {
  if (all(Ae[i, 1:dim(A)[2]]==0)) {
    if (Ae[i,dim(Ae)[2]]==0) {
      msg <- "There are infinitely many solutions."
    } else {
      msg <- "There are no solutions."
    }
    onesolution <- FALSE
  }
}
```

**SOLUTION:**

```{r}
if (onesolution) {
  msg <- "There is only one solution."
  Ae[,dim(Ae)[2]]
}

print(msg)
```

#### EXAMPLE 4

```{r}
# Coefficient matrix
A <- matrix(c(1,-3,4,-2,6,-8,1,-3,4), nrow = 3)
A

# Vector of constants
b <- c(3,-9,12)
b

# Row-equivalent matrix in echelon form
Ae <- echelon(A, b)
Ae

# Find and output solution(s)
onesolution <- TRUE
for(i in 1:dim(A)[1]) {
  if (all(Ae[i, 1:dim(A)[2]]==0)) {
    if (Ae[i,dim(Ae)[2]]==0) {
      msg <- "There are infinitely many solutions."
    } else {
      msg <- "There are no solutions."
    }
    onesolution <- FALSE
  }
}
```

**SOLUTION:**

```{r}
if (onesolution) {
  msg <- "There is only one solution."
  Ae[,dim(Ae)[2]]
}

print(msg)
```
