---
title: "DATA 605 Week 11 Homework"
author: "Ilya Kats"
date: "November 12, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

### Task

Using the `cars` dataset in R, build a linear model for stopping distance as a function of speed and
replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and
residual analysis).

### Analysis

Dataset `cars` includes 50 observations with 2 variables - `dist` containing stoppping distance in feet and `speed` containing speed of a car before applying the brakes in miles per hour. Data were recorded in the 1920s.

```{r}
summary(cars)
```

Let us look at the plot of two variables. Speed is the **explanatory** variable and stopping distance is the **response** one.

```{r}
plot(cars$speed, cars$dist, xlab='Speed (mph)', ylab='Stopping Distance (ft)', 
     main='Stopping Distance vs. Speed')
```

Let us build a linear model and find the best fitting line. 

```{r}
cars_lm <- lm(cars$dist ~ cars$speed)
cars_lm

plot(cars$speed, cars$dist, xlab='Speed (mph)', ylab='Stopping Distance (ft)', 
     main='Stopping Distance vs. Speed')
abline(cars_lm)
```

There appears to be some correlation between two variables, but let us evaluate the linear model we have. 

```{r}
summary(cars_lm)
```

The median value of the residuals is somewhat close to zero and quartiles and min/max values are roughly the same magnitude. The standard error of the `speed` variable is more than 9 times smaller than the corresponding coefficient. There should not be a lot of variability in this coefficient. On the other hand, the difference between the intercept estimate and standard error is less significant, so there may be more variability. The `speed` coefficient is highly significant. The intercept coefficient is less significant, but it is still relevant depending on the confidence interval desired. Finally, $R^2$ explains about 65.11% of the data's variation.

```{r}
plot(cars_lm$fitted.values, cars_lm$residuals, xlab='Fitted Values', ylab='Residuals')
abline(0,0)
```

It is possible to say that the outlier values do not show the same variance of the residuals; however, it is not very clear. I think it is reasonable to continue with the analysis and assume similar variance of residuals.

```{r}
qqnorm(cars_lm$residuals)
qqline(cars_lm$residuals)
```

Althought again there are some problems at the outlier levels, the normal Q-Q plot of the residuals appears to follow the theoretical line. Residuals are reasonably normally distributed.

### Conclusion

I believe the linear model does a good job at explaining the data. There appears to be some slight curvature in the main plot and in the residuals plot, so I decided to try a simple quadratic model (see below). It has it's own problems - again varability of residuals is not constant enough, q-q plot has some deviations, coefficients are not very significant and $R^2$ is not increased by much. I don't think it's an improvement over a simplier linear model.

#### Quadtratic Model

```{r}
speed <- cars$speed
speed2 <- speed^2
dist <- cars$dist
  
cars_qm <- lm(dist ~ speed + speed2)
summary(cars_qm)

speedvalues <- seq(0, 25, 0.1)
predictedcounts <- predict(cars_qm,list(speed=speedvalues, speed2=speedvalues^2))

plot(speed, dist, pch=16, xlab='Speed (mph)', ylab='Stopping Distance (ft)')
lines(speedvalues, predictedcounts)

plot(cars_qm$fitted.values, cars_qm$residuals, xlab='Fitted Values', ylab='Residuals')
abline(0,0)

qqnorm(cars_qm$residuals)
qqline(cars_qm$residuals)
```