---
title: "DATA 605 Week 3 Homework"
author: "Ilya Kats"
date: "September 17, 2017"
output:
  pdf_document: default
  html_document: default
params:
  password: ''
  username: root
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Set 1

### Part (1)

```{r}
A <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow=4, byrow=TRUE)

# Consider matrix A
A

# RREF of A
library(pracma)
rref(A)
```

There are 4 pivot rows, so the rank of $A$ is 4. Double-check with R.

```{r}
# Rank of A
Rank(A)
```

### Part (2)

Since rank of a matrix is defined as the number of linearly independent column vectors and is equal to the number of linearly independent row vectors, given an $m\times n$ matrix with $m>n$, the maximum rank is the lower value $rank_{max}=n$. Assuming a non-zero matrix, the rank should be at least 1, so $rank_{min}=1$.

### Part (3)

```{r}
B <- matrix(c(1,2,1,3,6,3,2,4,2), nrow=3, byrow=TRUE)

# Consider matrix B
B

# Rank of B
Rank(B)
```

In addition to getting an answer in R, we can see that row 2 and row 3 are multiples of row 1, so rows 1 and 2 and 1 and 3 are linearly dependent. There is only 1 linearly independent row, so $rank(B)=1$.

## Problem Set 2

```{r}
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow=3, byrow=TRUE)

# Consider matrix A
A
```

Since $A$ is a **triangular matrix**, its eigenvalues are values on the diagonal, so $\lambda_1=1$, $\lambda_2=4$ and $\lambda_3=6$.

```{r}
# Double-check eigenvalues in R
eigen(A)$values
```

The **characteristic polynomial** is 
$p_A(\lambda) = (1-\lambda)(4-\lambda)(6-\lambda)$ or
$p_A(\lambda) = 24-34\lambda+11\lambda^2-\lambda^3$.

If $\lambda=1$, then $A - 1I_3$ is row-reduced to 

```{r}
rref(A - 1 * diag(3))
```

$$
 \begin{bmatrix}
 0 &1 &0\\
 0 &0 &1\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$
Then $v_1=v_1$ and $v_2=0$ and $v_3=0$. The **eigenspace** is  
$$
 E_{\lambda=1}=
 \Bigg\langle \Bigg\{
 \begin{bmatrix}
 1\\
 0\\
 0
 \end{bmatrix}
 \Bigg\} \Bigg \rangle
$$

If $\lambda=4$, then $A - 4I_3$ is row-reduced to 

```{r}
rref(A - 4 * diag(3))
```

$$
 \begin{bmatrix}
 1 &-\frac{2}{3} &0\\
 0 &0 &1\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$

Then $v_1 - \frac{2}{3}v_2=0$ and $v_3=0$. 

Or $v_1=v_1$ and $v_2=\frac{3}{2}v_1=1.5v_1$ and $v_3=0$. 

The **eigenspace** is  

$$
 E_{\lambda=4}=
 \Bigg\langle \Bigg\{
 \begin{bmatrix}
 1\\
 1.5\\
 0
 \end{bmatrix}
 \Bigg\} \Bigg \rangle
$$

Finally, if $\lambda=6$, then $A - 6I_3$ is row-reduced to 

```{r}
rref(A - 6 * diag(3))
```

$$
 \begin{bmatrix}
 1 &0 &-1.6\\
 0 &1 &-2.5\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$

Then $v_1-1.6v_3=0$ and $v_2-2.5v_3=0$. 

Or $v_1=1.6v_3$ and $v_2=2.5v_3$ and $v_3=v_3$. 

The **eigenspace** is  

$$
 E_{\lambda=6}=
 \Bigg\langle \Bigg\{
 \begin{bmatrix}
 1.6\\
 2.5\\
 1
 \end{bmatrix}
 \Bigg\} \Bigg \rangle
$$