---
title: "DATA 605 Week 12 Discussion"
author: "Ilya Kats"
date: "November 18, 2017"
params:
  username: "root"
  password: ""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

## Multiple Regression Model

### Data

For this discussion, I figured some courses in Statistics will have sample data sets that will fit the requirements. I found an interesting one at the site for Princeton's course in Generalized Linear Models (http://data.princeton.edu/wws509/datasets/#salary).

I am using salary data that include observations on six variables for 52 tenure-track professors in a small college. The variables are:

- `sx`: Sex, coded 1 for female and 0 for male
- `rk`: Rank, coded 1 for assistant professor, 2 for associate professor, and 3 for full professor
- `yr`: Number of years in current rank
- `dg`: Highest degree, coded 1 if doctorate, 0 if masters
- `yd`: Number of years since highest degree was earned
- `sl`: Academic year salary, in dollars.

```{r}
# Data import
library(foreign)
salary <- read.dta("http://data.princeton.edu/wws509/datasets/salary.dta")

# Data size
dim(salary)

# General data description
summary(salary)

# Data sample
knitr::kable(head(salary))
```

Convert *sex* and *rank* into numerical representation. 

```{r}
salary$sx <- as.character(salary$sx)
salary$sx[salary$sx == "Male"] <- 0
salary$sx[salary$sx == "Female"] <- 1
salary$sx <- as.integer(salary$sx)

salary$rk <- as.character(salary$rk)
salary$rk[salary$rk == "Assistant"] <- 1
salary$rk[salary$rk == "Associate"] <- 2
salary$rk[salary$rk == "Full"] <- 3
salary$rk <- as.integer(salary$rk)
```

### Linear Model

I would like to build a model to predict professor's salary. All variables seem like they may be relevant.

Sex (`sx`) will be my **dichotomous** variable. 

It may make sense that increase in rank does not have a strictly linear relationship with the target variable. Higher rank may warrant higher salary increases, so my model will use square of rank (`rk`$^2$) as a **quadratic** variable.

Finally, I wanted to see if the model is influenced by **interaction** between sex and number of years since highest degree was earned (`sx`$\times$`yd`).

```{r}
# Quadratic variable
rk2 <- salary$rk^2

# Dichotomous vs. quantative interaction
sx_yd <- salary$sx * salary$yd

# Initial model
salary_lm <- lm(sl ~ sx + rk + rk2 + yr + dg + yd + sx_yd, data=salary)
summary(salary_lm)
```

Perform **backwards elimination** - removing one variable (the one with highest p-value) at a time. Removing *sex*.

```{r}
# Version 2
salary_lm <- update(salary_lm, .~. -sx)
summary(salary_lm)
```

Removing *square of rank*.

```{r}
# Version 3
salary_lm <- update(salary_lm, .~. -rk2)
summary(salary_lm)
```

Removing *highest degree*.

```{r}
# Version 4
salary_lm <- update(salary_lm, .~. -dg)
summary(salary_lm)
```

Removing *interaction between sex and number of years since highest degree was earned*.

```{r}
# Version 5
salary_lm <- update(salary_lm, .~. -sx_yd)
summary(salary_lm)
```

Removing *number of years since highest degree was earned*.

```{r}
# Version 6
salary_lm <- update(salary_lm, .~. -yd)
summary(salary_lm)
```

The final model has two variables - *rank* and *number of years in current rank* - that can be used to predict the target variable.

The median of residuals is a little off zero and they are a little uneven (the maximum value is noticeably farther away than the minimum value). Standard error for the `yr` variable could be smaller. 

Two coefficients imply that for every increase in rank the salary increases by $4,731.26 and with every year in the current rank the salary increases by $376.50.

Based on the Residuals vs. Fitted plot below there are some outliers in the data, but overall variability is fairly consistent. Based on the Q-Q plot, distribution of residuals is close to normal.

Based on $R^2$ value, the model explains 84.36% of variability in the data. 

```{r}
plot(salary_lm$fitted.values, salary_lm$residuals, xlab="Fitted Values", ylab="Residuals", main="Residuals vs. Fitted")
abline(h=0)

qqnorm(salary_lm$residuals)
qqline(salary_lm$residuals)
```
