---
title: "DATA 605 Week 15 Discussion"
author: "Ilya Kats"
date: "December 6, 2017"
params:
  username: "root"
  password: ""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(knitr)
library(rgl)
opts_chunk$set(echo = TRUE, error = TRUE)
knit_hooks$set(webgl = hook_webgl)
```

## Chapter 12 Section 1 Exercise 17

Describe in words and sketch the level curves for the given c values. 

$$
f(x,y) = x-y^2; c=-2, 0, 2
$$

### Solution

Consider $c=0$. Then $f(x,y) = x-y^2 = c = 0$, so $y = \pm\sqrt{x}$.

The horizontal plane at $c=0$ will intersect the surface following a parabola that is facing right. 

With $c=-2$, $y=\pm\sqrt{x+2}$ and with $c=2$, $y=\pm\sqrt{x-2}$, so the intersection will have the same shape (parabola) just shifted by 2. 

Let us plot the 3 level curves. 

```{r}
library(ggplot2)

x <- seq(-5,20,0.05)
y <- seq(-5, 5,0.05)
xy <- data.frame(expand.grid(x=x, y=y))
z <- xy$x-xy$y^2
f <- data.frame(xy, z)
ggplot(f, aes(x, y, z=z))+
  geom_raster(aes(fill=z))+
  geom_contour(breaks=c(-2,0,2), colour="white")+
  xlab("x")+ylab("y")
```

Even with contour coloring, it is somewhat difficult to imagine the surface. Luckily, `rgl` makes it incredible easy. And you can use your mouse to adjust graph perspective. Embedding the graph into an RMarkdown file proved to be more difficult.

```{r result, webgl=TRUE, results='hide'}
myColorRamp <- function(colors, values) {
    v <- (values - min(values))/diff(range(values))
    x <- colorRamp(colors)(v)
    rgb(x[,1], x[,2], x[,3], maxColorValue = 255)
}
cols <- myColorRamp(c("red", "blue"), f$y) 
plot3d(f$x, f$y, f$z, xlab="x", ylab="y", zlab="", col = cols)
```
